Learning action models with minimal observability

### 2018-12-09 doc/aij18/major-revision/background.tex
Sections 1 and 2 updated

This section serves two purposes: first we introduce basic planning concepts as well as the classical planning model that we will use throughout the rest of the paper; and secondly, we summarize the existing approaches to learn classical planning action models and highlight our contributions comparing with related work.

@@

We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state. A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e. either~$l=f$ or $l=\neg f$. A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not assign conflicting values to any fluent). We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.

A {\em state} $s$ is a full assignment of values to fluents and we explicitly include negative literals $\neg f$ in states; i.e. $|s|=|F|$, so the size of the state space is $2^{|F|}$. Like in PDDL~\cite{fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$. Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$. Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.

%simplifies subsequent definitions but often, we will abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as it is common in \strips\ planning.

A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. An action $a\in A$ is defined with:
\begin{itemize}
\item $\pre(a)\in\mathcal{L}(F)$, the {\em preconditions} of $a$, is the set of literals that must hold for the action $a\in A$ to be applicable.
\item $\eff^+(a)\in\mathcal{L}(F)$, the {\em positive effects} of $a$, is the set of literals that are true after the application of the action $a\in A$.
\item $\eff^-(a)\in\mathcal{L}(F)$, the {\em negative effects} of $a$, is the set of literals that are false after the application of the action.
\end{itemize}
We assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$ and that actions $a\in A$ are instantiated from given action schemas, as in PDDL. We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$. The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.

@@

Our approach copes with the partial observability of the plan execution adopting the \emph{open world assumption}. With regard to the observed states in a plan trace, we say that $\tup{s_0, s_1, \ldots, s_n}$ is a fully-observable (\FO) state trajectory if every state $s_i$ is a full assignment of values to fluents and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$ is composed of a single action; that is, $\theta(s_i,\tup{a})=s_{i+1}$. Otherwise, it is a partially-observable (\PO) state trajectory, meaning that at least one state $s_i$ is a partial assignment of values to fluents in which one or more literals are missing (formally, $|s_i|<|F|$). This implies that in a \PO state trajectory all fluents of a state $s_i$ might be missing, in which case, $s_i$ is a \emph{missing} or \emph{empty} state. This general definition of \PO gives rise to two particular cases:

\begin{itemize}
\item when \textbf{all} the $n-1$ intermediate states of a trajectory $s$ are \textbf{missing}, $s$ is a \emph{non-observable} (\NO) state trajectory.
\item when \textbf{none} of the $n-1$ intermediate states of a trajectory $s$ is \textbf{missing}, we will refer to $s$ as a \POstar state trajectory.
\end{itemize}

@@

\begin{table}[hbt!]
\centering
\begin{tabular}{c|c|c|}
	     & {\bf \# intermediate states} & {\bf state type} \\ \hline
    \FO & $n-1$  & {\small $\forall i, 1 \leq i < n$}  \\  & & $s_i$ is a full assignment \\ \hline
    \multirow{1}{*}{\POstar} & $n-1$ & {\small $\exists i, 1 \leq i < n$}  \\ & & $s_i$ is a partial assignment \\ \hline
    \multirow{1}{*}{\PO} & $\leq n-1$ & {\small $\exists i, 1 \leq i < n$}   \\  & & $s_i$ is a partial assignment \\ \hline
    \NO & 0 & {\small $\forall i, 1 \leq i < n$}  \\  & & $|s_i|=0$
\end{tabular}
\caption{Classification of state trajectories accordingly to the observed information.}
\label{tab:state_trajectory}
\end{table}

With regard to the observed actions in a plan trace, we say that $\tup{a_1, \ldots, a_n}$ is a \FO action sequence if it contains all the necessary actions to transit every state $s_{i-1}$ to the corresponding successor state $s_{i}$, for each {\small $1\leq i\le n$};  we say it is a \PO action sequence if at least one of these necessary actions is missing in the sequence, and a \NO action sequence when the sequence of observed actions is empty.

A plan trace $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$ for a planning frame $\Phi=\tup{F,A}$ holds that $a_i\in A$ for every action in $\tau$ and that $s_i \in \mathcal{L}(F)$ for each {\small $1\leq i\le n$}. Plan traces can be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).

@@

In this section we summarize the most recent and relevant approaches to learning action models found in the literature. Approaches will be examined according to the following parameters: the input knowledge (plan traces) accepted by the system, the expressiveness of the learned action model and the principal technique used for learning the action model (Table \ref{table:models_comparison1}), as well as the characteristics of the evaluation method to validate the learned models (Table \ref{table:models_comparison2}).

@@

The task of learning from less constrained traces subsumes learning from more constrained ones. Consequently, approaches to learning from, for instance traces with \PO state trajectories, will also be able to learn from traces with \POstar state trajectories. All the approaches analyzed in this work accept the more constrained definition of partial observations of intermediate states \POstar, and most of them also allow the sequence of intermediate states to be empty. Exceptionally, a \NO state sequence in \LOCM is a fully-empty trajectory, with neither initial or final state.

@@

\begin{table}
	\small
	\centering
	\begin{tabular}{ l | c | c | c }
		& \multicolumn{1}{c|}{\bf Input plan traces}
        & \multicolumn{1}{c|}{\bf Learned action model}
        & \multicolumn{1}{c}{\bf Technique}     \\
		\hline			
		\multirow{2}{*}{\ARMS} & \NO states & \strips & MAX-SAT \\ & \FO actions & & \\
        \hline
        \multirow{2}{*}{\SLAF} & \POstar states  & universal quantifiers in $\eff$ & logical inference \\ & \FO actions &  & SAT solver \\
         \hline
        %\multirow{2}{*}{\SLAF} & PO states and FO actions & existential quantifiers in $\pre$ \\ & & universal quantifiers in $\eff$ &  bayesian learning \\ & & & logical filtering \\
         %\hline
		\multirow{2}{*}{\LAMP} & \PO states &  quantifiers &  Markov logic networks \\  & \FO actions & logical implications &  \\
         \hline
         \AMAN & \NO states & \strips & graphical model estimation \\ & noisy actions & & \\
         \hline
         \NOISTA & \POstar and noisy states & \strips &  classification \\ & \FO actions & & \strips \texttt{} rules derivation \\
         \hline
         \CAMA & \PO states &  \strips & crowdsourcing annotation\\ & \FO actions &  & MAX-SAT \\
         \hline
         \LOCMtwo & \NO states &  predicates and types & Finite State Machines \\ & \FO actions & & \\
         \hline
		\FAMA & \NO states & \strips &   compilation to planning\\ & \NO actions & & \\
         \hline
	\end{tabular}
	\caption{Characteristics of action-model learning approaches}
	\label{table:models_comparison1}
\end{table}

@@

\begin{table}
	\small
	\centering
	\begin{tabular}{ l | c | c | c }
		%& \multicolumn{1}{c}{Evaluation task}
        & \textbf{Evaluation method} & \textbf{Metrics} & \textbf{\#tested domains/}   \\
        &   &   & \textbf{training data size} \\
		\hline			
		\multirow{1}{*}{\ARMS} & cross-validation with a test set & error counting of \#$\pre$ satisfaction  & 6  \\
        & of plan traces & and redundancy & 1,600-4,320 actions\\ & & & (160 plan traces) \\
        \hline
         \SLAF &  manual checking wrt GTM &  ---   & 4\\ & & & 1,000 actions\\
         \hline
		\multirow{1}{*}{\LAMP} & checking wrt GTM  & error counting of extra & 4\\ & & and missing \#$\pre$ and \#$\eff$ & 1,300-6,100 actions\\
           & & & (100-200 plan traces) \\
         \hline
         \AMAN & checking wrt GTM &  error counting of extra &  3 \\ & & and missing \#$\pre$ and \#$\eff$ & 40-200 plan traces\\
         \hline
         \NOISTA & checking wrt GTM  & error counting of extra & 5\\ & & and missing \#$\pre$ and \#$\eff$ & 5,000-20,000 actions\\
         \hline
         \CAMA & checking wrt GTM  &  error counting of extra & 3 \\ & & and missing \#$\pre$ and \#$\eff$ & 15-75 plan traces\\
         \hline
		\LOCMtwo & manual checking wrt GTM &  ---  &   --- \\
         \hline
		\FAMA & checking wrt GTM  & precision and recall & 15\\ & &  & 25 actions\\
         \hline
	\end{tabular}
	\caption{Evaluation of action models (GTM: ground-truth model)}
	\label{table:models_comparison2}
\end{table}

@@

The Learning Object-Centred Models (\textbf{\LOCM}) is an approach that only requires the \FO action sequence as input knowledge, without need for providing any information about the predicates or the state trajectory, not even the initial or final state~\cite{CresswellMW09,cresswell2013acquiring}. The lack of available state information is overcome by exploiting assumptions about the structure of the actions. Particularly, \LOCM assumes that objects found in the same position in the header of actions are grouped as a collection of objects (sorts) whose defined set of states is captured by a parameterized Finite State Machine (FSM). The intuitive assumptions of \LOCM, like the continuity of object transitions or the association of parameters between consecutive actions in the training dataset, yield a learning model heavily reliant on the kind of domain structure. A later work, \textbf{\LOCMtwo}, extends the applicability of the \LOCM algorithm to a wider range of domains by introducing a richer representation that allows using multiple FSMs to represent the state of a sort~\cite{cresswell2011generalised}.

@@

>>>>

This section serves two purposes; first, we introduce basic planning concepts and define the classical planning model we aim to learn; secondly, we summarize the most relevant existing approaches to learn classical planning action models.

@@

We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state. A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e. either~$l=f$ or $l=\neg f$. A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not assign conflicting values to any fluent). \textcolor[rgb]{1.00,0.00,0.00}{The complement of $L$ is defined as $\neg L=\{\neg l:l\in L\}$.} We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.

\textcolor[rgb]{1.00,0.00,0.00}{In \FAMA, we need to implicitly represent the unobserved literals of states. In order to do so, we will adopt the \emph{open world assumption}; i.e., what is not known to be true in a state is unknown. Consequently, states will explicitly include positive literals ($f$) and negative literals ($\neg f$) such that literals that are not in a state are unknown or unobserved.} Hence, a {\em state} $s$ is a full assignment of values to fluents; i.e. $|s|=|F|$, so the size of the state space is $2^{|F|}$. Like in PDDL~\cite{fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$. Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$. Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.

A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. An action $a\in A$ has a set of preconditions $\pre(a)\in\mathcal{L}(F)$ and a set of effects $\eff(a)\in\mathcal{L}(F)$. An action $a\in A$ is applicable in a given state $s$ iff $pre(a)\subseteq s$, i.e.~if the literals $pre(a)$ hold in $s$. \textcolor[rgb]{1.00,0.00,0.00}{The result of executing an applicable action $a\in A$ in a state $s$ is a new state $\theta(s,a)=\{s\setminus \neg\eff(a)\cup\eff(a)\}$. Note that subtracting the complement of $\eff(a)$ from $s$ ensures that $\theta(s,a)$ remains a well-defined state with positive and negative literals. Then:}

\begin{itemize}
\item \textcolor[rgb]{1.00,0.00,0.00}{$\eff^+(a)\in\mathcal{L}(F)$ is the {\em positive effects} of $a$, the subset of action effects that assert a positive literal in the state resulting after the application of $a$}
\item \textcolor[rgb]{1.00,0.00,0.00}{$\eff^-(a)\in\mathcal{L}(F)$ is the {\em negative effects} of $a$, the subset of action effects that assert a negative literal in the state resulting after the application of $a$}
\end{itemize}

\textcolor[rgb]{1.00,0.00,0.00}{Since we restrict our attention to \strips\ action models learning, we will assume the set of syntactic constraints imposed by \strips\ models, namely }that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$. Additionally, actions $a\in A$ are instantiated from given action schemas, as in PDDL.

%We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$. The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.

@@

Our approach copes with partial observability in the plan traces. Let $\pi=\tup{a_1, \ldots, a_n}$ be the plan executed by an agent that induces the state trajectory $s=\tup{s_0, s_1, \ldots, s_n}$, and let $\tau = \langle s_0, \ldots, a_i, \ldots, s_j, \ldots, s_n \rangle$ be the plan trace observed from the plan execution. With regards to the observed states of $\tau$, that we will refer to as $\tau_s$, we identify two general cases of observability:

\begin{enumerate}
\item we say that $\tau_s$ is a fully-observable (\FO) state trajectory if every observed intermediate state of $\tau_s$ is a full assignment of values to fluents, and \textcolor[rgb]{1.00,0.00,0.00}{there exists a single action that transitions from every state $s_i$ to state $s_{i+1}$ in $\tau_s$};  that is $\theta(s_i,\tup{a})=s_{i+1}$. \textcolor[rgb]{1.00,0.00,0.00}{This case clearly states that $\tau_s = s$, meaning that $\forall s_i \in \tau_s$, $s_i$ comprises all the literals of the corresponding state in the trajectory $s$ of the plan $\pi$. Formally, $\forall i, 1 \leq i  < n, |s_i| = |F|$.}
\item we say that $\tau_s$ is a partially-observable (\PO) state trajectory if at least one intermediate state of $\tau_s$ is a partial assignment of values to fluents. \textcolor[rgb]{1.00,0.00,0.00}{Formally, $\exists i, 1 \leq i  < n, |s_i| < |F|$.} This means that one or more literals are missing in the intermediate $s_i$, all of which may be missing. When all literals are missing, $s_i$ is a \emph{missing} or \emph{empty state} ($s_i = \emptyset$).
\end{enumerate}

The general definition of a \PO state trajectory gives rise to two special cases:

\begin{enumerate}
\item when \textbf{all} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, $\tau_s$ is a \emph{non-observable} (\NO) state trajectory. \textcolor[rgb]{1.00,0.00,0.00}{Formally, $\forall i, 1 \leq i < n,  s_i= \emptyset$; i.e., $|s_i| = 0$.}
\item when \textbf{none} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, we will refer to $\tau_s$ as a \POstar state trajectory. \textcolor[rgb]{1.00,0.00,0.00}{Formally, $\exists i, 1 \leq i < n, s_i \neq \emptyset$; i.e., $0 < |s_i| < |F|$.}
\end{enumerate}

@@

\begin{table}[hbt!]
\centering
\begin{tabular}{c|c|c|}
	     & {\bf \# intermediate states} & {\bf state type} \\ \hline
    \FO & $n-1$  & {\small $\forall i, 1 \leq i < n$}  \\  & & $s_i$ is a full assignment \textcolor[rgb]{1.00,0.00,0.00}{$|s_i|=|F|$} \\ \hline
    \multirow{1}{*}{\POstar} & $n-1$ & {\small $\exists i, 1 \leq i < n$}  \\ & & $s_i$ is a partial assignment \textcolor[rgb]{1.00,0.00,0.00}{$0 < |s_i|< |F|$}\\ \hline
    \multirow{1}{*}{\PO} & $\leq n-1$ & {\small $\exists i, 1 \leq i < n$}   \\  & & $s_i$ is a partial assignment \textcolor[rgb]{1.00,0.00,0.00}{$|s_i|< |F|$}\\ \hline
    \NO & 0 & {\small $\forall i, 1 \leq i < n$}  \\  & & $s_i$ \textcolor[rgb]{1.00,0.00,0.00}{is an empty state  $|s_i|=0$}
\end{tabular}
\caption{Classification of state trajectories accordingly to the observed information.}
\label{tab:state_trajectory}
\end{table}

\textcolor[rgb]{1.00,0.00,0.00}{\FAMA can also deal with partial observability in the observed actions of $\tau$, that we will refer to as $\tau_a$. We identify three levels of observability, from the greatest to the lowest:}

\begin{enumerate}
\item \textcolor[rgb]{1.00,0.00,0.00}{when \textbf{all} of the actions of $\pi$ appear in $\tau_a$, we say that $\tau_a$ is a fully-observable (\FO) action sequence; i.e.,  $\tau_a=\pi$. In this case, $\tau_a$ contains all the necessary actions to transit every state $s_{i-1}$ to its corresponding successor state $s_{i}$, from $s_0$ to $s_n$. This is the type of input trace accepted by all the existing learning approaches (see section \ref{related_work} for details)}.
\item \textcolor[rgb]{1.00,0.00,0.00}{when \textbf{some} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a partially observable (\PO) action sequence. In this case, at least one of the necessary actions of the plan $\pi$ is missing in $\tau_a$. Formally, $\exists i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$.}
\item \textcolor[rgb]{1.00,0.00,0.00}{when \textbf{none} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a non-observable (\NO) action sequence. Formally, $\forall i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$. That is, $\tau_a = \emptyset$.}
\end{enumerate}

================

A plan trace $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$ for a planning frame $\Phi=\tup{F,A}$ holds that $a_i\in A$ for every action in $\tau$ and that $s_i \in \mathcal{L}(F)$ for each {\small $1\leq i\le n$}.

NO TENGO CLARO LO QUE QUEREMOS DECIR CON ESTA FRASE (??). ADEMAS, NO ES CORRECTA PORQUE LOS ESTADOS DE UNA TRAZA PUEDEN SER FULL-OBSERVABLE Y $\mathcal{L}(F)$ SON PARTIAL ASSIGNMENTS OF VALUES TO FLUENTS.

=================

Plan traces can be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO). \textcolor[rgb]{1.00,0.00,0.00}{In section \ref{overview}, we expose the impact of all combinations of observed state trajectories and observed action sequences in the complexity of the learning task.}

@@

In this section we summarize the most recent and relevant approaches to learning action models found in the literature. Approaches will be examined according to the following parameters: the observability of the plan traces accepted by the system, the expressiveness of the learned action model and the principal technique used for learning the action model (Table \ref{table:models_comparison1}), as well as the characteristics of the evaluation method used to validate the learned models (Table \ref{table:models_comparison2}).

@@

The task of learning from less constrained traces subsumes learning from more constrained ones. Consequently, approaches to learning from, say traces with \PO state trajectories, will also enable learning from traces with \POstar state trajectories. All the approaches analyzed in this work accept the more constrained definition of partial observations of intermediate states \POstar, two of them also allow the sequence of intermediate states to be empty (\PO) and the majority accept \NO state trajectories. \textcolor[rgb]{1.00,0.00,0.00}{Exceptionally, \LOCM is the only approach capable of learning from a fully-empty state trajectory, with neither initial nor final state.}

@@

\begin{table}
	\small
	\centering
	\begin{tabular}{ l | c | c | c }
		& \multicolumn{1}{c|}{\bf Input plan traces}
        & \multicolumn{1}{c|}{\bf Learned action model}
        & \multicolumn{1}{c}{\bf Technique}     \\
		\hline			
		\multirow{2}{*}{\ARMS} & \NO states & \strips & MAX-SAT \\ & \FO actions & & \\
        \hline
        \multirow{2}{*}{\SLAF} & \POstar states  & universal quantifiers in $\eff$ & logical inference \\ & \FO actions &  & SAT solver \\
         \hline
        %\multirow{2}{*}{\SLAF} & PO states and FO actions & existential quantifiers in $\pre$ \\ & & universal quantifiers in $\eff$ &  bayesian learning \\ & & & logical filtering \\
         %\hline
		\multirow{2}{*}{\LAMP} & \PO states &  quantifiers &  Markov logic networks \\  & \FO actions & logical implications &  \\
         \hline
         \AMAN & \NO states & \strips & graphical model estimation \\ & noisy actions & & \\
         \hline
         \NOISTA & \POstar and noisy states & \strips &  classification \\ & \FO actions & & \strips \texttt{} rules derivation \\
         \hline
         \CAMA & \PO states &  \strips & crowdsourcing annotation\\ & \FO actions &  & MAX-SAT \\
         \hline
         \textcolor[rgb]{1.00,0.00,0.00}{\LOUGA} & \textcolor[rgb]{1.00,0.00,0.00}{\NO states} & \textcolor[rgb]{1.00,0.00,0.00}{\strips} & \textcolor[rgb]{1.00,0.00,0.00}{Genetic algorithm} \\ & \textcolor[rgb]{1.00,0.00,0.00}{\FO actions} & \textcolor[rgb]{1.00,0.00,0.00}{negative preconditions} & \\
         \hline
         \LOCMtwo & \textcolor[rgb]{1.00,0.00,0.00}{--- }&  predicates and types & Finite State Machines \\ & \FO actions & & \\
         \hline
		\FAMA & \NO states & \strips &   compilation to planning\\ & \NO actions & & \\
         \hline
	\end{tabular}
	\caption{Characteristics of action-model learning approaches}
	\label{table:models_comparison1}
\end{table}

@@

\begin{table}
	\small
	\centering
	\begin{tabular}{ l | c | c | c }
		%& \multicolumn{1}{c}{Evaluation task}
        & \textbf{Evaluation method} & \textbf{Metrics} & \textbf{\#tested domains/}   \\
        &   &   & \textbf{training data size} \\
		\hline			
		\multirow{1}{*}{\ARMS} & cross-validation with a test set & error counting of \#$\pre$ satisfaction  & 6  \\
        & of plan traces & and redundancy & 1,600-4,320 actions\\ & & & (160 plan traces) \\
        \hline
         \SLAF &  manual checking wrt GTM &  ---   & 4\\ & & & 1,000 actions\\
         \hline
		\multirow{1}{*}{\LAMP} & checking wrt GTM  & error counting of extra & 4\\ & & and missing \#$\pre$ and \#$\eff$ & 1,300-6,100 actions\\
           & & & (100-200 plan traces) \\
         \hline
         \AMAN & checking wrt GTM &  error counting of extra &  3 \\ & & and missing \#$\pre$ and \#$\eff$ & 40-200 plan traces\\
         \hline
         \NOISTA & checking wrt GTM  & error counting of extra & 5\\ & & and missing \#$\pre$ and \#$\eff$ & 5,000-20,000 actions\\
         \hline
         \CAMA & checking wrt GTM  &  error counting of extra & 3 \\ & & and missing \#$\pre$ and \#$\eff$ & 15-75 plan traces\\
         \hline
       \textcolor{red}{\LOUGA} & \textcolor[rgb]{1.00,0.00,0.00}{cross-validation with} & \textcolor[rgb]{1.00,0.00,0.00}{redundant effects} & \textcolor[rgb]{1.00,0.00,0.00}{5} \\  & \textcolor[rgb]{1.00,0.00,0.00}{a test set of plan traces} & \textcolor[rgb]{1.00,0.00,0.00}{differences wrt the test set} & \textcolor[rgb]{1.00,0.00,0.00}{800 - 3200 actions (160 traces)}\\
         \hline
		\LOCMtwo & manual checking wrt GTM &  ---  &   --- \\
         \hline
		\FAMA & checking wrt GTM  & precision and recall & 14\\  & validation with a test set &  & 20-50 actions\\
         \hline
	\end{tabular}
	\caption{Evaluation of action models (GTM: ground-truth model)}
	\label{table:models_comparison2}
\end{table}

@@

\textcolor[rgb]{1.00,0.00,0.00}{One of the latest incorporations to the family of action model learning algorithms is \LOUGA. This system uses a genetic algorithm to learn the effects of actions. In order to do this, each gene in the genome encodes whether a predicate is a positive effect, negative effect or none for a particular action, and the fitness of an individual is evaluated by reproducing the trace with the model encoded in the individual. After a solution for the effects is found, an ad-hoc algorithm is used to infer preconditions by finding those literals that are always present before the execution of an action. \LOUGA evaluates the learned models via cross-validation using the same metrics they use in their fitness function. In more detail, they measure (1) redundant positive and negative effects, (2) preconditions not met, and (3) literals observed in the input trace but not in the corresponding execution of the plan with the learned model.}

\vspace{0.15cm}

\textcolor[rgb]{1.00,0.00,0.00}{Finally, we present the Learning Object-Centred Models (\textbf{\LOCM}), possibly the most distinctive learning system due to its ability of learning with minimal input knowledge.} \textbf{\LOCM} only requires the \FO action sequence of the plan trace, without need for providing any information about the predicates or the state trajectory, not even the initial or final state~\cite{CresswellMW09,cresswell2013acquiring}. The lack of available state information is overcome by exploiting assumptions about the structure of the actions. Particularly, \LOCM assumes that objects found in the same position in the header of actions are grouped as a collection of objects (sorts) whose defined set of states is captured by a parameterized Finite State Machine (FSM). The intuitive assumptions of \LOCM, like the continuity of object transitions or the association of parameters between consecutive actions in the training dataset, yield a learning model heavily reliant on the kind of domain structure. A later work, \textbf{\LOCMtwo}, extends the applicability of the \LOCM algorithm to a wider range of domains by introducing a richer representation that allows using multiple FSMs to represent the state of a sort~\cite{cresswell2011generalised}.

@@

\textcolor[rgb]{1.00,0.00,0.00}{The distinctive feature of \LOCMtwo lies in the need to learn the state variables (fluents) because predicates are not either provided as input nor they are deducible from the plan trace as no state observability is allowed. In contrast, \FAMA and the rest of approaches either assume the set of predicates are provided alongside the input traces or assume they are extractable from the observed states of the plan trace, in which case the plan trace must comprise at least a grounded sample of every predicate. Similarly, the syntax of an action header (the action name and its parameters) is either extractable from the action sequence of the plan trace or it must be explicitly provided.}
